---
title: "Evaluate Louvain performance"
author: "Marc Jaskir"
date: "1/9/2023"
output: 
  html_document: 
    toc: yes
---

# Load packages/data
```{r}
rm(list=ls())

library(ggcorrplot)
library(ggplot2)
library(igraph)
library(visNetwork)
library(DescTools)
library(mgcv)
library(gam)

louvain_observed <- read.csv('/Users/mjaskir/ngg/rotations/alexanderbloch/neuromaps/outputs/2/observed/louvain_performance.csv')
louvain_null <- read.csv('/Users/mjaskir/ngg/rotations/alexanderbloch/neuromaps/outputs/2/null/louvain_performance.csv')
```

# Visualize Louvain performance by resolution

## Fit models predicting NMI, ARI, and #Communities by Resolution
```{r}
### NMI ###

# Observed
louvain_observed_nmi_loess <- loess(nmi ~ res, data=louvain_observed, span=0.05)
louvain_observed_nmi_loess_fit <- data.frame(res=sort(louvain_observed$res), nmi=sort(louvain_observed$nmi), nmi_predicted=sort(predict(louvain_observed_nmi_loess)))

# Null
louvain_null_nmi_loess <- loess(nmi ~ res, data=louvain_null, span=0.05)
louvain_null_nmi_loess_fit <- data.frame(res=sort(louvain_null$res), nmi=sort(louvain_null$nmi), nmi_predicted=sort(predict(louvain_null_nmi_loess)))

# Difference in fits
fit_differences <- c()
for (resolutions in unique(louvain_observed_nmi_loess_fit$res)) {
  fit_observed_mean <- mean(louvain_observed_nmi_loess_fit[louvain_observed_nmi_loess_fit$res == resolutions,'nmi_predicted'], na.rm=TRUE)
  fit_null_mean <- mean(louvain_null_nmi_loess_fit[louvain_null_nmi_loess_fit$res == resolutions,'nmi_predicted'], na.rm = TRUE)
  fit_differences <- c(fit_differences, fit_observed_mean - fit_null_mean)
}
louvain_diff_nmi_loess_fit <- data.frame(res=unique(louvain_observed_nmi_loess_fit$res), nmi_diff=fit_differences)

### ARI ###

# Observed
louvain_observed_ari_loess <- loess(ari ~ res, data=louvain_observed, span=0.05)
louvain_observed_ari_loess_fit <- data.frame(res=sort(louvain_observed$res), ari=sort(louvain_observed$ari), ari_predicted=sort(predict(louvain_observed_ari_loess)))

# Null
louvain_null_ari_loess <- loess(ari ~ res, data=louvain_null, span=0.05)
louvain_null_ari_loess_fit <- data.frame(res=sort(louvain_null$res), ari=sort(louvain_null$ari), ari_predicted=sort(predict(louvain_null_ari_loess)))

# Difference in fits
fit_differences <- c()
for (resolutions in unique(louvain_observed_ari_loess_fit$res)) {
  fit_observed_mean <- mean(louvain_observed_ari_loess_fit[louvain_observed_ari_loess_fit$res == resolutions,'ari_predicted'], na.rm=TRUE)
  fit_null_mean <- mean(louvain_null_ari_loess_fit[louvain_null_ari_loess_fit$res == resolutions,'ari_predicted'], na.rm = TRUE)
  fit_differences <- c(fit_differences, fit_observed_mean - fit_null_mean)
}
louvain_diff_ari_loess_fit <- data.frame(res=unique(louvain_observed_ari_loess_fit$res), ari_diff=fit_differences)

### Number of communities ###

# Observed
louvain_observed_num_communities_loess <- loess(num_communities ~ res, data=louvain_observed, span=0.05)
louvain_observed_num_communities_loess_fit <- data.frame(res=sort(louvain_observed$res), num_communities=sort(louvain_observed$num_communities), num_communities_predicted=sort(predict(louvain_observed_num_communities_loess)))

# Null
louvain_null_num_communities_loess <- loess(num_communities ~ res, data=louvain_null, span=0.05)
louvain_null_num_communities_loess_fit <- data.frame(res=sort(louvain_null$res), num_communities=sort(louvain_null$num_communities), num_communities_predicted=sort(predict(louvain_null_num_communities_loess)))

# Difference in fits
fit_differences <- c()
for (resolutions in unique(louvain_observed_num_communities_loess_fit$res)) {
  fit_observed_mean <- mean(louvain_observed_num_communities_loess_fit[louvain_observed_num_communities_loess_fit$res == resolutions,'num_communities_predicted'], na.rm=TRUE)
  fit_null_mean <- mean(louvain_null_num_communities_loess_fit[louvain_null_num_communities_loess_fit$res == resolutions,'num_communities_predicted'], na.rm = TRUE)
  fit_differences <- c(fit_differences, fit_observed_mean - fit_null_mean)
}
louvain_diff_num_communities_loess_fit <- data.frame(res=unique(louvain_observed_num_communities_loess_fit$res), num_communities_diff=fit_differences)

```

## Plot NMI by Resolution
```{r}
par(mfrow = c(2,2))

plot(x=louvain_observed$res, y=louvain_observed$nmi, cex=0.05, ylim=c(0,0.75), col='grey', xlab=expression(gamma), ylab='NMI', main='Observed', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_observed_nmi_loess_fit$res, y=louvain_observed_nmi_loess_fit$nmi_predicted, col="red")
plot(x=louvain_null$res, y=louvain_null$nmi, cex=0.05, ylim=c(0,0.75), col='grey', xlab=expression(gamma), ylab='NMI', main='Null', cex.lab=1.25, cex.main=1.25)
  lines(x=louvain_null_nmi_loess_fit$res, y=louvain_null_nmi_loess_fit$nmi_predicted)
plot(x=louvain_observed_nmi_loess_fit$res, y=louvain_observed_nmi_loess_fit$nmi_predicted, ylim=c(0,0.75), col="red", type='l', xlab=expression(gamma), ylab='NMI', main='Overlay', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_null_nmi_loess_fit$res, y=louvain_null_nmi_loess_fit$nmi_predicted)
plot(x=louvain_diff_nmi_loess_fit$res, y=louvain_diff_nmi_loess_fit$nmi_diff, col="blue", type='l', xlab=expression(gamma), ylab='Difference', main='Observed - Null', cex.lab=1.25, cex.main=1.25)
```

## Plot ARI by Resolution
```{r}
par(mfrow = c(2,2))

plot(x=louvain_observed$res, y=louvain_observed$ari, cex=0.05, col='grey', ylim=c(0,0.6), xlab=expression(gamma), ylab='ARI', main='Observed', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_observed_ari_loess_fit$res, y=louvain_observed_ari_loess_fit$ari_predicted, col="red")
plot(x=louvain_null$res, y=louvain_null$ari, cex=0.05, col='grey', ylim=c(0,0.6), xlab=expression(gamma), ylab='ARI', main='Null', cex.lab=1.25, cex.main=1.25)
  lines(x=louvain_null_ari_loess_fit$res, y=louvain_null_ari_loess_fit$ari_predicted)
plot(x=louvain_observed_ari_loess_fit$res, y=louvain_observed_ari_loess_fit$ari_predicted, col="red", type='l', ylim=c(0,0.6), xlab=expression(gamma), ylab='ARI', main='Overlay', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_null_ari_loess_fit$res, y=louvain_null_ari_loess_fit$ari_predicted)
plot(x=louvain_diff_ari_loess_fit$res, y=louvain_diff_ari_loess_fit$ari_diff, col="blue", type='l', xlab=expression(gamma), ylab='Difference', main='Observed - Null', cex.lab=1.25, cex.main=1.25)
```

## Plot #Communities by Resolution
```{r}
par(mfrow = c(2,2))

plot(x=louvain_observed$res, y=louvain_observed$num_communities, cex=0.05, col='grey', xlab=expression(gamma), ylab='Number of communities', main='Observed') +
  lines(x=louvain_observed_num_communities_loess_fit$res, y=louvain_observed_num_communities_loess_fit$num_communities_predicted, col="red")
plot(x=louvain_null$res, y=louvain_null$num_communities, cex=0.05, col='grey', xlab=expression(gamma), ylab='Number of communities', main='Null')
  lines(x=louvain_null_num_communities_loess_fit$res, y=louvain_null_num_communities_loess_fit$num_communities_predicted)
plot(x=louvain_observed_num_communities_loess_fit$res, y=louvain_observed_num_communities_loess_fit$num_communities_predicted, col="red", type='l', xlab=expression(gamma), ylab='Number of communities', main='Overlay') +
  lines(x=louvain_null_num_communities_loess_fit$res, y=louvain_null_num_communities_loess_fit$num_communities_predicted)
plot(x=louvain_diff_num_communities_loess_fit$res, y=louvain_diff_num_communities_loess_fit$num_communities_diff, col="blue", type='l', xlab=expression(gamma), ylab='Difference', main='Observed - Null', cex.lab=1.25, cex.main=1.25)
```
## Determine optimal resolution as maximal difference between observed and null NMI
```{r}
louvain_diff_nmi_loess_fit_optimal <- louvain_diff_nmi_loess_fit[louvain_diff_nmi_loess_fit$nmi_diff == max(louvain_diff_nmi_loess_fit$nmi_diff, na.rm=TRUE),]
louvain_diff_nmi_loess_fit_optimal <- louvain_diff_nmi_loess_fit_optimal[! is.na(louvain_diff_nmi_loess_fit_optimal$nmi_diff),]
print(louvain_diff_nmi_loess_fit_optimal)
```
Optimal Louvain resolution: 0.83

# Visualize Louvain performance by #Communities

## Fit models predicting NMI and ARI by #Communities
```{r}
### NMI ###

# Observed
louvain_observed_nmi_loess <- loess(nmi ~ num_communities, data=louvain_observed, span=0.2, control=loess.control(surface="direct"))
louvain_observed_nmi_loess_fit <- data.frame(num_communities=sort(louvain_observed$num_communities), nmi=sort(louvain_observed$nmi), nmi_predicted=sort(predict(louvain_observed_nmi_loess)))

# Null
louvain_null_nmi_loess <- loess(nmi ~ num_communities, data=louvain_null, span=0.2, control=loess.control(surface="direct"))
louvain_null_nmi_loess_fit <- data.frame(num_communities=sort(louvain_null$num_communities), nmi=sort(louvain_null$nmi), nmi_predicted=sort(predict(louvain_null_nmi_loess)))

# Difference in fits
fit_differences <- c()
for (num in unique(louvain_observed_nmi_loess_fit$num_communities)) {
  fit_observed_mean <- mean(louvain_observed_nmi_loess_fit[louvain_observed_nmi_loess_fit$num_communities == num,'nmi_predicted'], na.rm=TRUE)
  fit_null_mean <- mean(louvain_null_nmi_loess_fit[louvain_null_nmi_loess_fit$num_communities == num,'nmi_predicted'], na.rm = TRUE)
  fit_differences <- c(fit_differences, fit_observed_mean - fit_null_mean)
}
louvain_diff_nmi_loess_fit <- data.frame(num_communities=unique(louvain_observed_nmi_loess_fit$num_communities), nmi_diff=fit_differences)

### ARI ###

# Observed
louvain_observed_ari_loess <- loess(ari ~ num_communities, data=louvain_observed, span=0.2, control=loess.control(surface="direct"))
louvain_observed_ari_loess_fit <- data.frame(num_communities=sort(louvain_observed$num_communities), ari=sort(louvain_observed$ari), ari_predicted=sort(predict(louvain_observed_ari_loess)))

# Null
louvain_null_ari_loess <- loess(ari ~ num_communities, data=louvain_null, span=0.2, control=loess.control(surface="direct"))
louvain_null_ari_loess_fit <- data.frame(num_communities=sort(louvain_null$num_communities), ari=sort(louvain_null$ari), ari_predicted=sort(predict(louvain_null_ari_loess)))

# Difference in fits
fit_differences <- c()
for (num_communitiesolutions in unique(louvain_observed_ari_loess_fit$num_communities)) {
  fit_observed_mean <- mean(louvain_observed_ari_loess_fit[louvain_observed_ari_loess_fit$num_communities == num_communitiesolutions,'ari_predicted'], na.rm=TRUE)
  fit_null_mean <- mean(louvain_null_ari_loess_fit[louvain_null_ari_loess_fit$num_communities == num_communitiesolutions,'ari_predicted'], na.rm = TRUE)
  fit_differences <- c(fit_differences, fit_observed_mean - fit_null_mean)
}
louvain_diff_ari_loess_fit <- data.frame(num_communities=unique(louvain_observed_ari_loess_fit$num_communities), ari_diff=fit_differences)

```

## Plot NMI by #Communities
```{r}
par(mfrow = c(2,2))

plot(x=louvain_observed$num_communities, y=louvain_observed$nmi, cex=0.05, ylim=c(0,0.75), col='grey', xlab='Number of communities', ylab='NMI', main='Observed', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_observed_nmi_loess_fit$num_communities, y=louvain_observed_nmi_loess_fit$nmi_predicted, col="red")
plot(x=louvain_null$num_communities, y=louvain_null$nmi, cex=0.05, ylim=c(0,0.75), col='grey', xlab='Number of communities', ylab='NMI', main='Null', cex.lab=1.25, cex.main=1.25)
  lines(x=louvain_null_nmi_loess_fit$num_communities, y=louvain_null_nmi_loess_fit$nmi_predicted)
plot(x=louvain_observed_nmi_loess_fit$num_communities, y=louvain_observed_nmi_loess_fit$nmi_predicted, col="red", ylim=c(0,0.75), type='l', xlab='Number of communities', ylab='NMI', main='Overlay', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_null_nmi_loess_fit$num_communities, y=louvain_null_nmi_loess_fit$nmi_predicted)
plot(x=louvain_diff_nmi_loess_fit$num_communities, y=louvain_diff_nmi_loess_fit$nmi_diff, xlim=c(0,50),col="blue", type='l', xlab='Number of communities', ylab='Difference', main='Observed - Null', cex.lab=1.25, cex.main=1.25)
```

## Plot ARI by #Communities
```{r}
par(mfrow = c(2,2))

plot(x=louvain_observed$num_communities, y=louvain_observed$ari, cex=0.05, col='grey', xlab='Number of communities', ylab='ARI', main='Observed', cex.lab=1.25, cex.main=1.25) +
  lines(x=louvain_observed_ari_loess_fit$num_communities, y=louvain_observed_ari_loess_fit$ari_predicted, col="red")
plot(x=louvain_null$num_communities, y=louvain_null$ari, cex=0.05, col='grey', xlab='Number of communities', ylab='ARI', main='Null', cex.lab=1.25, cex.main=1.25)
  lines(x=louvain_null_ari_loess_fit$num_communities, y=louvain_null_ari_loess_fit$ari_predicted)
plot(x=louvain_observed_ari_loess_fit$num_communities, y=louvain_observed_ari_loess_fit$ari_predicted, col="red", type='l', xlab='Number of communities', ylab='ARI', main='Overlay', cex.lab=1.25, cex.main=1.25, ylim=c(0.05,0.5)) +
  lines(x=louvain_null_ari_loess_fit$num_communities, y=louvain_null_ari_loess_fit$ari_predicted)
plot(x=louvain_diff_ari_loess_fit$num_communities, y=louvain_diff_ari_loess_fit$ari_diff, col="blue", type='l', xlab='Number of communities', ylab='Difference', main='Observed - Null', cex.lab=1.25, cex.main=1.25)
```

# Session info
```{r}
sessionInfo()
```
