---
title: "Visualize community-level spatial averages"
author: "Marc Jaskir"
date: "1/9/2023"
output: 
  html_document: 
    toc: yes
---

# Load packages/data
```{r}
library(cifti)
library(ciftiTools)
library(gifti)

rm(list=ls())

# Load community assignments
community_assignment <- read.table('/Users/mjaskir/ngg/rotations/alexanderbloch/neuromaps/outputs/4/fsLR_32k_adjacency_matrix_modZ_norm_community_assignment.csv',sep=',')
colnames(community_assignment) <- c('map','community')

# Load reference maps (transformed to fsLR 32k)
setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/neuromaps/outputs/1/annotations_fslr')
map_files <- list.files(pattern="*.gii")
map_labels <- gsub('.gii','',map_files)
map_list <- lapply(map_files, function(x){read_gifti(x)})
names(map_list) <- map_labels

# Load surface files
setwd('/Users/mjaskir/neuromaps-data/atlases/fsLR')
fslr_32k_midthickness_lh <- read_gifti('tpl-fsLR_den-32k_hemi-L_midthickness.surf.gii')
fslr_32k_midthickness_rh <- read_gifti('tpl-fsLR_den-32k_hemi-R_midthickness.surf.gii')

# Load medial wall files
setwd('/Users/mjaskir/ngg/rotations/alexanderbloch/neuromaps/scripts/medial_wall')
medial_wall_lh <- read.table('medialwall.mask.leftcortex.csv',sep=',')
medial_wall_rh <- read.table('medialwall.mask.rightcortex.csv',sep=',')
```

# Rescale reference maps
Each map is rescaled to have range 0-1
NOTE: vertex data were labeled with shape, normal, or estimate
```{r}
map_list_rescaled <- list()
for (label in map_labels) {
  
  if (length(map_list[label][[1]]$data$shape) != 0) {
    
    vertex_data <- map_list[label][[1]]$data$shape
    
    # Make minimum 0, if necessary
    if (range(vertex_data,na.rm = TRUE)[1] != 0) {
      vertex_data <- vertex_data - range(vertex_data,na.rm=TRUE)[1]
    }

    # Rescale to 0-1
    vertex_data <- vertex_data/range(vertex_data,na.rm=TRUE)[2]
    
    # Remove medial wall
    if (grepl('_lh$',label)) {
      vertex_data <- vertex_data*medial_wall_lh
    } else {
      vertex_data <- vertex_data*medial_wall_rh
    }
    
    # Save to list
    map_list_rescaled[label] <- map_list[label]
    map_list_rescaled[label][[1]]$data$shape <- vertex_data
    
  } else if (length(map_list[label][[1]]$data$normal) != 0) {
    
    vertex_data <- map_list[label][[1]]$data$normal

    # Make minimum 0, if necessary
    if (range(vertex_data,na.rm = TRUE)[1] != 0) {
      vertex_data <- vertex_data - range(vertex_data,na.rm=TRUE)[1]
    }

    # Rescale to 0-1
    vertex_data <- vertex_data/range(vertex_data,na.rm=TRUE)[2]
    
    # Remove medial wall
    if (grepl('_lh$',label)) {
      vertex_data <- vertex_data*medial_wall_lh
    } else {
      vertex_data <- vertex_data*medial_wall_rh
    }
    
    # Save to list
    map_list_rescaled[label] <- map_list[label]
    map_list_rescaled[label][[1]]$data$normal <- vertex_data
    
  } else if (length(map_list[label][[1]]$data$estimate) != 0) {
    
    vertex_data <- map_list[label][[1]]$data$estimate
    
    # Make minimum 0, if necessary
    if (range(vertex_data,na.rm = TRUE)[1] != 0) {
      vertex_data <- vertex_data - range(vertex_data,na.rm=TRUE)[1]
    }

    # Rescale to 0-1
    vertex_data <- vertex_data/range(vertex_data,na.rm=TRUE)[2]
    
    # Remove medial wall
    if (grepl('_lh$',label)) {
      vertex_data <- vertex_data*medial_wall_lh
    } else {
      vertex_data <- vertex_data*medial_wall_rh
    }
    
    # Save to list
    map_list_rescaled[label] <- map_list[label]
    map_list_rescaled[label][[1]]$data$estimate <- vertex_data
    
  } else {
    print('WARNING: Unusual data structure')
  }
  
}
```

# Parse reference maps by community assignment
```{r}
# Divide by community assignment
community_1 <- list()
community_2 <- list()
community_3 <- list()
for (map in community_assignment$map) {
  community <- community_assignment[community_assignment$map == map,'community']
  for (label in grep(map,map_labels)) {
    if (community == 1) {
      community_1[map_labels[label]] <- map_list_rescaled[label]
    } else if (community == 2) {
      community_2[map_labels[label]] <- map_list_rescaled[label]
    } else if (community == 3) {
      community_3[map_labels[label]] <- map_list_rescaled[label]
    }
  }
}

# Divide further by hemisphere
community_1_lh <- community_1[grep('lh',names(community_1))]
community_1_rh <- community_1[grep('rh',names(community_1))]
community_2_lh <- community_2[grep('lh',names(community_2))]
community_2_rh <- community_2[grep('rh',names(community_2))]
community_3_lh <- community_3[grep('lh',names(community_3))]
community_3_rh <- community_3[grep('rh',names(community_3))]
```


# Create community-specific spatial average surfaces
Average spatial maps are rescaled to have range 0-1
```{r}
# Create template surface
community_avg_template <- map_list[1]
names(community_avg_template) <- NULL

# Compute average
n <- length(community_1_lh)
for (label in names(community_1_lh)) {
  if (label == names(community_1_lh)[1]) {
    vertex_data <- as.vector(unlist(community_1_lh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_1_lh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_1_lh_avg <- community_avg_template
community_1_lh_avg[[1]]$data$shape <- vertex_data_avg

# Compute average
n <- length(community_1_rh)
for (label in names(community_1_rh)) {
  if (label == names(community_1_rh)[1]) {
    vertex_data <- as.vector(unlist(community_1_rh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_1_rh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_1_rh_avg <- community_avg_template
community_1_rh_avg[[1]]$data$shape <- vertex_data_avg

# Compute average
n <- length(community_2_lh)
for (label in names(community_2_lh)) {
  if (label == names(community_2_lh)[1]) {
    vertex_data <- as.vector(unlist(community_2_lh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_2_lh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_2_lh_avg <- community_avg_template
community_2_lh_avg[[1]]$data$shape <- vertex_data_avg

# Compute average
n <- length(community_2_rh)
for (label in names(community_2_rh)) {
  if (label == names(community_2_rh)[1]) {
    vertex_data <- as.vector(unlist(community_2_rh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_2_rh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_2_rh_avg <- community_avg_template
community_2_rh_avg[[1]]$data$shape <- vertex_data_avg

# Compute average
n <- length(community_3_lh)
for (label in names(community_3_lh)) {
  if (label == names(community_3_lh)[1]) {
    vertex_data <- as.vector(unlist(community_3_lh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_3_lh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_3_lh_avg <- community_avg_template
community_3_lh_avg[[1]]$data$shape <- vertex_data_avg

# Compute average
n <- length(community_3_rh)
for (label in names(community_3_rh)) {
  if (label == names(community_3_rh)[1]) {
    vertex_data <- as.vector(unlist(community_3_rh[label][[1]]$data))
  } else {
    vertex_data <- vertex_data + as.vector(unlist(community_3_rh[label][[1]]$data))
  }
}
vertex_data_avg <- vertex_data/n

# Rescale to 0-1
vertex_data_avg <- vertex_data_avg - range(vertex_data_avg,na.rm=TRUE)[1]
vertex_data_avg <- vertex_data_avg/range(vertex_data_avg,na.rm=TRUE)[2]

# Save to template surface
community_3_rh_avg <- community_avg_template
community_3_rh_avg[[1]]$data$shape <- vertex_data_avg
```

# Create community-specific spatial maxima surfaces
```{r}
community_1_lh_max <- community_avg_template
community_1_lh_max[[1]]$data$shape <- ifelse(community_1_lh_avg[[1]]$data$shape - community_2_lh_avg[[1]]$data$shape > 0, ifelse(community_1_lh_avg[[1]]$data$shape - community_3_lh_avg[[1]]$data$shape > 0,1,0),0)

community_1_rh_max <- community_avg_template
community_1_rh_max[[1]]$data$shape <- ifelse(community_1_rh_avg[[1]]$data$shape - community_2_rh_avg[[1]]$data$shape > 0, ifelse(community_1_rh_avg[[1]]$data$shape - community_3_rh_avg[[1]]$data$shape > 0,1,0),0)

community_2_lh_max <- community_avg_template
community_2_lh_max[[1]]$data$shape <- ifelse(community_2_lh_avg[[1]]$data$shape - community_1_lh_avg[[1]]$data$shape > 0, ifelse(community_2_lh_avg[[1]]$data$shape - community_3_lh_avg[[1]]$data$shape > 0,1,0),0)

community_2_rh_max <- community_avg_template
community_2_rh_max[[1]]$data$shape <- ifelse(community_2_rh_avg[[1]]$data$shape - community_1_rh_avg[[1]]$data$shape > 0, ifelse(community_2_rh_avg[[1]]$data$shape - community_3_rh_avg[[1]]$data$shape > 0,1,0),0)

community_3_lh_max <- community_avg_template
community_3_lh_max[[1]]$data$shape <- ifelse(community_3_lh_avg[[1]]$data$shape - community_1_lh_avg[[1]]$data$shape > 0, ifelse(community_3_lh_avg[[1]]$data$shape - community_2_lh_avg[[1]]$data$shape > 0,1,0),0)

community_3_rh_max <- community_avg_template
community_3_rh_max[[1]]$data$shape <- ifelse(community_3_rh_avg[[1]]$data$shape - community_1_rh_avg[[1]]$data$shape > 0, ifelse(community_3_rh_avg[[1]]$data$shape - community_2_rh_avg[[1]]$data$shape > 0,1,0),0)
```

# Visualize spatial averages for each community
```{r}
community_1.cifti <- as_cifti(cortexL = community_1_lh_avg[[1]]$data$shape, cortexR = community_1_rh_avg[[1]]$data$shape)
community_1.cifti <- add_surf(community_1.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)
community_2.cifti <- as_cifti(cortexL = community_2_lh_avg[[1]]$data$shape, cortexR = community_2_rh_avg[[1]]$data$shape)
community_2.cifti <- add_surf(community_2.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)
community_3.cifti <- as_cifti(cortexL = community_3_lh_avg[[1]]$data$shape, cortexR = community_3_rh_avg[[1]]$data$shape)
community_3.cifti <- add_surf(community_3.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)

view_cifti(community_1.cifti, widget=TRUE, zlim=c(0,1))
view_cifti(community_2.cifti, widget=TRUE, zlim=c(0,1))
view_cifti(community_3.cifti, widget=TRUE, zlim=c(0,1))
```

# Visualize spatial maxima for each community
```{r}
community_1.cifti <- as_cifti(cortexL = community_1_lh_max[[1]]$data$shape, cortexR = community_1_rh_max[[1]]$data$shape)
community_1.cifti <- add_surf(community_1.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)
community_2.cifti <- as_cifti(cortexL = community_2_lh_max[[1]]$data$shape, cortexR = community_2_rh_max[[1]]$data$shape)
community_2.cifti <- add_surf(community_2.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)
community_3.cifti <- as_cifti(cortexL = community_3_lh_max[[1]]$data$shape, cortexR = community_3_rh_max[[1]]$data$shape)
community_3.cifti <- add_surf(community_3.cifti, surfL = fslr_32k_midthickness_lh, surfR = fslr_32k_midthickness_rh)

view_cifti(community_1.cifti, widget=TRUE, zlim=c(0,1), colors=c("white","red"))
view_cifti(community_2.cifti, widget=TRUE, zlim=c(0,1), colors=c("white","forestgreen"))
view_cifti(community_3.cifti, widget=TRUE, zlim=c(0,1), colors=c("white","blue"))
```

# Session info
```{r}
sessionInfo()
```
